### 2020-04-01 05-translation.tex
major rewriting of examples
check carefully!

\[
\begin{array}{lcl@{\hskip 1cm}lcl}
 P_s &\defeq & P(s,\widetilde{s},\widehat{s},\underline{s}) 
 &
 \overline{P_s}& \defeq & P(\overline{s},\widetilde{s},\widehat{s},\underline{s})
\end{array}
\]
%{\color{red}
%\[
%\begin{array}{rcl}
% P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h_j,k_j\in\{0,1\}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j}\ \link{\hat{s}_i}{\hat{s}_o}.P_s\\
%&&\ + \\
%&&
%\sum_{\scriptsize \begin{array}{l}h_j,k_j, \\ h'_j\in\{0,1\}\end{array}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j} \  \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} \    (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h'_j} \ \link{\underline{s}_i}{\underline{s}_o}.P_s \\
%
%%\sum_{R_1+R_2\geq 0}(\sum_{h,k \in[0,1]} 
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_1} \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} (\sum_{h ,k\in[0,1]}
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_2} \link{\underline{s}_i}{\underline{s}_o}.P_s\\
%&&\ +\\
%&&\sum_{j\geq 0}(
%  \startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^j\ \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
%\end{array}
%\]
%}
\[
\begin{array}{rcl}
 P(s,\widetilde{s},\widehat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\widehat{s}_i}\chainedlink{\widehat{s}_o}{\noact}\chainend{\noact} (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ + \\
&&
\sum_{h\geq 0,k\geq 1}
 (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ +\\
&&\sum_{h\geq 0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
\end{array}
\]
The first line of $ P(s,\widetilde{s},\widehat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\widehat{s}_i}{\widehat{s}_o}$).
Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
combination is possible. By knowing the number of reactions that test $s$, we can bound the maximum 
values of $h$ and $k$.
The second line accounts for the analogous case where $s$ is not provided by the context ($\link{\underline{s}_i}{\underline{s}_o}$). 
The condition $k\geq 1$ guarantees that $s$ will remain present (the continuation is $P_s$).
%Here taking
%$h = \sum_{j\geq0} h_j +h'_j$, $k = (\sum_{j \geq 0} k_j )  + 1$ guarantees that $s$ will be present in the next step. Please note that 
The third line accounts for the case where $s$ is tested for presence, but it is neither produced nor provided by  the context. Therefore, in the next step $s$ will be absent in the system (the continuation is $\overline{P_s}$).
Note that in the case of $\overline{P_s}$ the test for presence of $s$ in the system is just replaced by the test for its absence.
%Process $P_s$ (and $\overline{P_s}$) offers link chain prefixes for all the possible
% permutations of the solid links; nevertheless if the reactions, and their enumeration, is known
% the code of $P_s$ can be drastically simplified.

@@

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
$$
\llbracket{\cal A},\gamma \rrbracket 
\triangleq
 \restrict{\mathit{names}}(I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1)
$$
where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}\, \cup\, \{cxt\}$. 
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
to allow the name $r_1$ to be matched at the start of any chain; also, 
for notational convenience, we fix that 
%$r_1 = \silent$,
$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the cardinality of $\mathit{reacts}$.
\end{definition}

@@

The topmost restriction $\restrict{\mathit{names}}$ appearing in the process $\llbracket{\cal A},\gamma\rrbracket$ serves to guarantee that all names appearing in a link of the chain labelling a transition are matched.
Since all names appearing in any prefix of $\llbracket{\cal A},\gamma\rrbracket$ are restricted, in the transition $\llbracket{\cal A},\gamma\rrbracket \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ it means that the observation $\restrict{\mathit{names}}\upsilon$ has the form $\link{\silent}{\silent}\dots\link{\silent}{\silent}$, i.e., it is silent, and that $\upsilon$ is solid. Later on we will be interested in reasoning about the actual chain $\upsilon$ used in the transition.  It has the peculiarity to start and end with silent actions and to include all names in $\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$. As a matter of notation we call such chain $\upsilon$ \emph{complete} (solid and with $\silent$ actions at the extremes). We will then use $\llparenthesis{\cal A},\gamma\rrparenthesis$ to refer to the encoding without topmost name restrictions, i.e.,
$$
\llparenthesis{\cal A},\gamma\rrparenthesis \defeq 
I~|~\Pi_{s\in C_0}P_{s} ~|~ \Pi_{s \notin C_0}\overline{P_s} ~|~ \Pi_{a\in A} P_a ~|~  \mathit{Cxt}^1
$$
and will write $P\xRightarrow{\upsilon} P'$ to mean that $P\xrightarrow{\upsilon} P'$ with $\upsilon$ complete.

@@

 \begin{restatable}[Correctness 1]{proposition}{propcorruno}
% \begin{proposition}(correctness 1)\\
 \label{prop:corr1}
% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
 Let 
 $P=\llbracket{\cal A},\gamma\rrbracket$ with
$$P  = \restrict{\mathit{names}} (I~|~\Pi_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \Pi_{s \in C_0}P_s ~|~ \Pi_{s \notin C_0}\overline{P}_s).
$$
 If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:
 \begin{enumerate}
 \item
  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$, and
  \item 
 $P'= \restrict{\mathit{names}} (I~|~\Pi_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$.
 \end{enumerate}
Moreover, given $\pi^1=(\gamma^1,\delta^1)$, we have $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
 %\end{proposition}
\end{restatable}

>>>>

\[
\begin{array}{lcl@{\hskip 1cm}lcl}
 P_s &\defeq & E(s,\widetilde{s},\widehat{s},\underline{s}) 
 &
 \overline{P_s}& \defeq & E(\overline{s},\widetilde{s},\widehat{s},\underline{s})
\end{array}
\]
%{\color{red}
%\[
%\begin{array}{rcl}
% P(s,\tilde{s},\hat{s},\underline{s}) & \defeq &\sum_{h_j,k_j\in\{0,1\}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j}\ \link{\hat{s}_i}{\hat{s}_o}.P_s\\
%&&\ + \\
%&&
%\sum_{\scriptsize \begin{array}{l}h_j,k_j, \\ h'_j\in\{0,1\}\end{array}}
%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h_j} \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^{k_j} \  \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} \    (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^{h'_j} \ \link{\underline{s}_i}{\underline{s}_o}.P_s \\
%
%%\sum_{R_1+R_2\geq 0}(\sum_{h,k \in[0,1]} 
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_1} \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact} (\sum_{h ,k\in[0,1]}
%%  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ (\startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact})^k)^{R_2} \link{\underline{s}_i}{\underline{s}_o}.P_s\\
%&&\ +\\
%&&\sum_{j\geq 0}(
%  \startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^j\ \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
%\end{array}
%\]
%}
\[
\begin{array}{rcl}
 E(s,\widetilde{s},\widehat{s},\underline{s}) & \defeq &\sum_{h,k\geq 0}  
  (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\widehat{s}_i}\chainedlink{\widehat{s}_o}{\noact}\chainend{\noact} (\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ + \\
&&
\sum_{h\geq 0,k\geq 1}
 (\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \startchain{\underline{s}_i}\chainedlink{\underline{s}_o}{\noact}\chainend{\noact}(\startchain{\widetilde{s}_i}\chainedlink{\widetilde{s}_o}{\noact}\chainend{\noact})^k.P_s\\
&&\ +\\
&&\sum_{h\geq 0}(\startchain{s_i}\chainedlink{s_o}{\noact}\chainend{\noact})^h \ \link{\underline{s}_i}{\underline{s}_o}.\overline{P_s}
\end{array}
\]
The first line of $ E(s,\widetilde{s},\widehat{s},\underline{s})$ accounts for the case where $s$ is tested for presence by $h$ reactions and produced by $k$ reactions, while being provided by the context ($\link{\widehat{s}_i}{\widehat{s}_o}$).
Thus, $s$ will be present at the next step (the continuation is $P_s$). Here $h$ and $k$ are not known a priori and therefore any
combination is possible. By knowing the number of reactions that test $s$, we can bound the maximum 
values of $h$ and $k$.
The second line accounts for the analogous case where $s$ is not provided by the context ($\link{\underline{s}_i}{\underline{s}_o}$). 
The condition $k\geq 1$ guarantees that $s$ will remain present (the continuation is $P_s$).
%Here taking
%$h = \sum_{j\geq0} h_j +h'_j$, $k = (\sum_{j \geq 0} k_j )  + 1$ guarantees that $s$ will be present in the next step. Please note that 
The third line accounts for the case where $s$ is tested for presence, but it is neither produced nor provided by  the context. Therefore, in the next step $s$ will be absent in the system (the continuation is $\overline{P_s}$).
Note that in the case of $\overline{P_s}$ the test for presence of $s$ in the system is just replaced by the test for its absence.
%Process $P_s$ (and $\overline{P_s}$) offers link chain prefixes for all the possible
% permutations of the solid links; nevertheless if the reactions, and their enumeration, is known
% the code of $P_s$ can be drastically simplified.

@@

\begin{definition}[Translation]
\label{def:trans}
Let ${\cal A} = (S, A)$ be a RS, and let $\pi=(\gamma,\delta)$ be an extended interactive process in ${\cal A}$, with  $\gamma=\{C_i\}_{i\in\mathbb{N}}$. 
We define its c\CNA~translation $\llbracket{\cal A},\gamma \rrbracket$ as follows: 
$$
\llbracket{\cal A},\gamma \rrbracket 
\triangleq
 \restrict{\mathit{names}}\left(I ~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1 ~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s}\right)
$$
where $\mathit{names} = \mathit{reacts}\, \cup\, \mathit{ents}\, \cup \, \mathit{prods}\, \cup\, \{cxt\}$. 
%For notational convenience, we fix that $r_1 = \silent$,  $r_{u+1} = cxt_1$ for $u$ the number of reacts, and $cxt_{w+1} = \silent$ for $w$ the number of entities.
For technical reasons, we introduce the init process $I \defeq \link{\silent}{r_1}.I$
to allow the name $r_1$ to be matched at the start of any chain; also, 
for notational convenience, we fix that 
%$r_1 = \silent$,
$r_{u+1} = \mathit{cxt}$ and $p_{u+1}=\silent$,  for $u$ the cardinality of $\mathit{reacts}$.
\end{definition}

@@

The topmost restriction $\restrict{\mathit{names}}$ appearing in the process $\llbracket{\cal A},\gamma\rrbracket$ serves to guarantee that all names appearing in a link of the chain labelling a transition are matched.
Since all names appearing in any prefix of $\llbracket{\cal A},\gamma\rrbracket$ are restricted, in the transition $\llbracket{\cal A},\gamma\rrbracket \xrightarrow{\restrict{\mathit{names}}\upsilon} P'$ it means that the observation $\restrict{\mathit{names}}\upsilon$ has the form $\link{\silent}{\silent}\dots\link{\silent}{\silent}$, i.e., it is silent, and that $\upsilon$ is solid. Later on we will be interested in reasoning about the actual chain $\upsilon$ used in the transition.  It has the peculiarity to start and end with silent actions and to include all names in $\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$. As a matter of notation we call such chain $\upsilon$ \emph{complete} (solid and with $\silent$ actions at the extremes). We will then use $\llparenthesis{\cal A},\gamma\rrparenthesis$ to refer to the encoding without topmost name restrictions, i.e.,
$$
\llparenthesis{\cal A},\gamma\rrparenthesis \defeq 
I~|~ \prod_{a\in A} P_a ~|~  \mathit{Cxt}^1~|~\prod_{s\in C_0}P_{s} ~|~ \prod_{s \notin C_0}\overline{P_s} 
$$
and will write $P\xRightarrow{\upsilon} P'$ to mean that $P\xrightarrow{\upsilon} P'$ with $\upsilon$ complete.

@@

 \begin{restatable}[Correctness 1]{proposition}{propcorruno}
% \begin{proposition}(correctness 1)\\
 \label{prop:corr1}
% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
 Let 
 $P=\llbracket{\cal A},\gamma\rrbracket$ with
$$P  = \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a ~|~ \mathit{Cxt}^1 ~|~ \prod_{s \in C_0}P_s ~|~ \prod_{s \notin C_0}\overline{P}_s\right).
$$
 If there exists  $P'$  such that $P \xrightarrow{v}P'$,  it holds that:
 \begin{enumerate}
 \item
  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$, and
  \item 
 $P'= \restrict{\mathit{names}} \left(I~|~\prod_{a \in A}P_a~|~ \mathit{Cxt}^2 ~|~ \prod_{s \in C_1 \cup D_1}P_s| \prod_{s \notin C_1 \cup D_1}\overline{P}_s\right)$.
 \end{enumerate}
Moreover, given $\pi^1=(\gamma^1,\delta^1)$, we have $P' = \llbracket {\cal A},\gamma^1\rrbracket$.
 %\end{proposition}
\end{restatable}
